---
title: "10.0.5"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 10.0.5

Loading data from an external source is a common task.  
In this section, we will look at some of the methods for loading data from a file.

## Loading from a database

The `DBI` package provides a common interface for connecting to databases. 
Let's use the `renv::install("DBI")` function to install the package.


The `DBI` package provides a common interface for connecting to databases.
We have a sqlite database in this project: `fake_data.`
Let's connect to the database.  We'll need to install the `RSQLite` package to connect to the database.

```{r}
# Function to check and install required packages
install_and_load <- function(packages) {
    for (package in packages) {
        if (!require(package, character.only = TRUE)) {
            renv::install(package, lock = TRUE)
        }
        library(package, character.only = TRUE)
    }
}

# List of required packages
packages <- c("DBI", "RSQLite", "jsonlite", "writexl", "xml2")

# Install and load missing packages
install_and_load(packages)

# Take a snapshot of the renv environment
renv::snapshot()
```

Once installed, we can connect to the database.

```{r}
# load USJudgeRatings data, convert the rowname  to a column
df <- USJudgeRatings
df <- cbind(rownames(df), data.frame(df, row.names=NULL))
# rename the new column
colnames(df)[1] <- "judge_name"
head(df)
```

```{r}
if (file.exists("judge_ratings.sqlite")) {
  file.remove("judge_ratings.sqlite")
}
# create a new database
dbname <- "judge_ratings.sqlite"
con <- dbConnect(RSQLite::SQLite(), dbname)
# create a table in the database
dbWriteTable(con, "judge_ratings", df, overwrite = TRUE)
# close the connection
dbDisconnect(con)
# query our new database
con <- dbConnect(RSQLite::SQLite(), dbname)
# query the database
query <- "SELECT * FROM judge_ratings"
# execute the query
result <- dbGetQuery(con, query)
# print the result
print(result)
# close the connection
dbDisconnect(con)
```


let's convert our data frame to a csv file.

```{r}
if (file.exists("judge_ratings.csv")) {
  file.remove("judge_ratings.csv")
}
write.csv(df, "judge_ratings.csv", row.names = FALSE)
```
and json

```{r}
if (file.exists("judge_ratings.json")) {
  file.remove("judge_ratings.json")
}
jsonlite::write_json(df, "judge_ratings.json")
```

and excel
```{r}
if (file.exists("judge_ratings.xlsx")) {
  file.remove("judge_ratings.xlsx")
}
writexl::write_xlsx(df, "judge_ratings.xlsx")
```



XML
```{r}
if (file.exists("judge_ratings.xml")) {
  file.remove("judge_ratings.xml")
}

#query our database, return the result as xml
con <- dbConnect(RSQLite::SQLite(), dbname)
# synthesize xml output via the sql query
query <- "SELECT '<judge_name>'||judge_name||'</judge_name> <CONT>'||CONT||'</CONT> <INTG>'||INTG||'</INTG> <DMNR>'||DMNR||'</DMNR> <DILG>'||DILG||'</DILG> <CFMG>'||CFMG||'</CFMG> <DECI>'||DECI||'</DECI> <PREP>'||PREP||'</PREP> <FAMI>'||FAMI||'</FAMI> <ORAL>'||ORAL||'</ORAL> <WRIT>'||WRIT||'</WRIT> <PHYS>'||PHYS||'</PHYS> <RTEN>'||RTEN||'</RTEN> <DECI>'||DECI||'</DECI> <PREP>'||PREP||'</PREP> <FAMI>'||FAMI||'</FAMI> <ORAL>'||ORAL||'</ORAL> <WRIT>'||WRIT||'</WRIT> <PHYS>'||PHYS||'</PHYS> <RTEN>'||RTEN||'</RTEN>' FROM judge_ratings"
result <- dbGetQuery(con, query)
# add the xml a root element of judge_ratings
# use the xml2 package to parse the xml
xml <- xml2::as_xml_document(paste0("<judge_ratings>", paste(result[,1], collapse = ""), "</judge_ratings>"))
# write the xml to a file
xml2::write_xml(xml, "judge_ratings.xml")
```

RDS
```{r}
if (file.exists("judge_ratings.rds")) {
  file.remove("judge_ratings.rds")
}
saveRDS(df, "judge_ratings.rds")
```

and Rdata
```{r}
rdata_df <- df
rm(df)
if (file.exists("judge_ratings.Rdata")) {
  file.remove("judge_ratings.Rdata")
}
save(df, file = "judge_ratings.Rdata")
```
