---
title: "5.1.1"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## T-tests

This notebook demonstrates the use of t-tests in R.

```{r}
# Load the data.csv file
data <- read.csv("data.csv")
# take the mean of `value2`
value2_mean <- mean(data$value2)
```

## One-Sample T-Test

The one-sample t-test is used to determine whether the mean of a single sample is significantly different from a known or hypothesized population mean.


```{r}
# Perform a one-sample t-test using the mean of `value2` against `value 1`
t.test(data$value1, mu = value2_mean)
```

## Independent T-Test

The independent t-test is used to determine whether the means of two independent groups are significantly different from one another.

```{r}
# Perform an independent t-test between `value1` and `value2`
t.test(data$value1, data$value2)
```

## Paired T-Test

The paired t-test is used to determine whether the means of two paired samples are significantly different from one another.

```{r}
# Perform a paired t-test between `value1` and `value2`
t.test(data$value1, data$value2, paired = TRUE)
```
